---
title: "Dow Jones Index Time Series Analysis"
output: html_notebook
---


Installing Important Packages
```{r}
install.packages("dplyr")
```

Importing the packages
```{r}
library(dplyr)
```

Importing Dow Jones Dataset
```{r}
index <- read.csv("./data/sp500_index.csv")
index
```

Checking the datatypes of all columns
```{r}
print(sapply(index, class))
```

Transforming the "Date" column from charater to Date
```{r}
transformed_index <- transform(index, Date = as.Date(Date))
```

Checking the Datatypes after transformation
```{r}
print(sapply(transformed_index, class))

```

Formatting the "Date" column to Year-month format inorder to group the index values
```{r}
transformed_index$Month_yr <- format(as.Date(transformed_index$Date), "%Y-%m")
```

```{r}
print(sapply(transformed_index, class))
```

```{r}
transformed_index
```

Averaging the index values for each month and creating a new DataFrame.
```{r}
avg_index <- transformed_index %>% group_by(Month_yr) %>% summarise(average_value = mean(S.P500))
```

```{r}
avg_index
```

Extracting the index values into a new list.
```{r}
index_values <- avg_index[[2]]
```

Plotting the index values
```{r}
plot(index_values)
```

Creating a Time Series Object from the index values.
```{r}
ts_index <- ts(index_values, start=c(2012, 12), end = c(2022, 12), frequency=12)
```

```{r}
ts_index
```

Plotting the Time Series Object.
```{r}
plot(ts_index)
```

